IFA Comparison of Noncognitive Skills \\ BCS and MCS

### 2018-07-10 codes/R/notes/backupcode.Rnw
move notes to separate folder

>>>>

% DECOMPOSITION GRAPHS AND TEXT

\clearpage
\subsection{Decomposition of inequality}

We have populations $k \in \mathcal{K}$. For each population, we observe covariates $X_k$. For some $j \subseteq \mathcal{K}$, we observe outcome $Y_j$. We can identify the covariate distributions $F_{X_k}$ and the conditional distributions $F_{Y_j \mid X_j}$ and quantile functions $Q_{Y_j \mid X_j}$. Define the support of $X_k$ as $\mathcal{X}_k$ and the region of interest for the outcome as $\mathcal{Y}_j$.

The interest is in the counterfactual distribution of the outcome:
$$ F_{Y(j \mid k)}(y) = \int_{\mathcal{X}_k} F_{Y_j \mid X_j}(y \mid x) d F_{X_k}(x), \qquad y \in \mathcal{Y}_j $$
and the corresponding quantiles:
$$ Q_{Y(j \mid k)}(\tau) = F^\leftarrow_{Y(j \mid k)}(\tau) $$

Given this, we can define counterfactual effects (CE):
\bi
\item Type 1 (Structure effect): CE of changing the conditional distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid k)}(\tau)$
\item Type 2 (Composition effect): CE of changing the covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(j \mid m)}(\tau)$
\item Type 3: CE of changing both conditional and covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid m)}(\tau)$
\ei

In our application, we have $k=BCS,MCS$. We can thus decompose the observed differences in the distribution of non-cognitive skills into structure and composition effects. For example, the difference in the distribution of externalising skills can be written as:
$$F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} =
  \left[ F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} \right] +
  \left[ F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} \right]$$

<<COUNTERF, dependson="FA_SCORES_MAIN">>=
# perform counterfactuals
@

<<COUNTERF_PREP_PLOTS, dependson='COUNTERF'>>=
# prepare counterfactual plots
@

\begin{landscape}
\begin{figure}
<<COUNTERF_EXT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.ext
@
\caption{Counterfactual decomposition -- externalising skills}\label{fig:counterf_ext}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
<<COUNTERF_INT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.int
@
\caption{Counterfactual decomposition -- internalising skills}\label{fig:counterf_int}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

<<COUNTERF_PREP_TAB, dependson='COUNTERF'>>=
# prepare counterfactual tables
@

<<COUNTERF_INEQ, dependson='COUNTERF_PREP_TABS', results='markup'>>=
cat("Males EXT")
print(iqdecomp(logitcf.ext.m))

cat("Males INT")
print(iqdecomp(logitcf.int.m))

cat("Females EXT")
print(iqdecomp(logitcf.ext.f))

cat("Females INT")
print(iqdecomp(logitcf.int.f))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% COUNTERFACTUAL DECOMPOSITIONS

<<COUNTERF_SEQDEC_ESTIM, dependson="SELECTVARS">>=
# estimate sequential decompositions
@

<<COUNTERF_SEQDEC_PREP>>=
# extract data
@

<<COUNTERF_SEQDEC_PLOT>>=
# decomposition plots
@

\begin{landscape}
\begin{figure}
<<COUNTERF_SEQDEC_EXT, dependson='COUNTERF_SEQDEC_PLOT', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.seqdec.ext
@
\caption{Counterfactual sequential decomposition -- externalising skills}\label{fig:counterf_seq_ext}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
<<COUNTERF_SEQDEC_INT, dependson='COUNTERF_SEQDEC_PLOT', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.seqdec.int
@
\caption{Counterfactual sequential decomposition -- internalising skills}\label{fig:counterf_seq_int}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
